---
title: "Lab 1"
output:
  html_document:
    toc: true
    toc_float: true
---

Libraries we will use (install if necessary):  

```{r eval = F}
install.packages("readr")
install.packages("igraph")
install.packages("dplyr")
install.packages("raster")
install.packages("scales")
install.packages("intergraph")
install.packages("network")
install.packages("RColorBrewer")
install.packages("tidygraph")
```



# 1. Network visualization

## 1.1 Loading the data and visualizing the network

```{r message = F, warning = F}
library(dplyr)
library(scales)
library(ggplot2)
library(ggraph)
# Libraries for network objects
library(igraph)
library(tidygraph)
# Loading the data from the STNet package
node <- STNet::SwinePrem
edge <- STNet::SwineMov
```

We use the function `graph_from_data_frame()`  to create the network from a `data.frame`. This function takes as arguments:`d`  which requires the contacts data, `vertices` which is the data of the nodes (farms in our example), and `directed` which specifies if we want to consider the direction of the contact.    

```{r}
# Create the network
net <- graph_from_data_frame(d=edge, vertices=node, directed=T)
# We can examine the contents of the network using the function summary()
summary(net)
```

The result will show a summary of the network. The first line shows that is an object class `igraph` with 40 nodes and 1611 contacts (edges).
Then we have a list of all the attributes of the network. The network attributes can be node attributes (n) or edge attributes (e).  
Here we can also see the attribute class:    
  - `c` = character.  
  - `n` = numeric.  
  
For example, the attribute "name(v/c)" its a node attribute of class character, and the attribute "pigs.moved (e/n)" is an attribute of the edge of class numeric.    

To visualize the network we can use the function `plot()`.  

```{r}
#plot network#
#for help look look for: ?plot.igraph#
plot(net)
```

This figure looks a bit messy, to make it more informative we can visualize dome of the parameters. We use the arguments `edge.arrow.size` to adjust the size of the arrowhead, `vertex.size` to adjust the node size, and `vertex.label` to remove the names of the nodes.  

```{r}
#make it clearer: adjust size, remove labels#
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.label=NA)
```

**Excercise:** Change the size of the nodes and the arrows to make a different graph.  


## 1.2 Node color

We can get specific attributes from the `igraph` object using `V()` for nodes, and `E()` for edges.  
The same way we do for a `data.frame`, we can create new attributes of the network using the `$` operator and assigning a name to a new variables. Now we will create a new variable for the node color.  

```{r}
# Create a color palette
colpal <- RColorBrewer::brewer.pal(length(levels(node$farm_type)), 
                                   "Dark2")
colpal <- rainbow(length(levels(node$farm_type)), s = 0.6)
# Assign a new variable for the color:
V(net)$color <- colpal[factor(V(net)$farm_type)]
#plot with colors
#vertex.color assigns colors by calling to the newly created "color" or "color2" variables that are now in the vertex (so node) dimension of the network#
#the vertex dimension is called to by naming it V(net)#
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.color=V(net)$color, vertex.label=NA)
legend("bottomleft", legend = levels(factor(V(net)$farm_type)), col = colpal, pch = 16, bty = "n")
```


## 1.3 Layouts

We can change the position of the nodes using the argument `layout`. Now we will use different layouts to see how the network changes.  

```{r}
#layout_randomly uses a random location
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.color=V(net)$color, vertex.label=NA, layout=layout_randomly)
legend("bottomleft", legend = levels(factor(V(net)$farm_type)), col = colpal, pch = 16, bty = "n")
```

Force directed layouts place the nodes in fixed places based on various criteria. The layouts `layout_nicely` and `layout_with_kk` force the network in clear layouts, based on the position of the nodes in the network in terms of connectivity.  They also attempt to keep the distance between connected nodes fixed.

```{r}
par(mfrow = c(1,2))
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.color=V(net)$color, vertex.label=NA, layout=layout_with_kk)
#the following layouts force nodes to adopt certain shapes#
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.color=V(net)$color, vertex.label=NA, layout=layout_in_circle)
```

**Excercise:** Try different layouts: `layout_nicely`, `layout_with_graphopt`, `layout_as_star`, `layout_with_fr`, `layout_on_grid`.    



## 1.4 Size of the edge

```{r}
# We will change the edge size based on the number of aimals moved
E(net)$width <- E(net)$pigs.moved
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.color=V(net)$color, vertex.label=NA, layout=layout_with_kk)
# Since the edgge sizes are very large, we will constrain the edge size based on a scale using the package scales
E(net)$width <- scales::rescale(E(net)$pigs.moved, c(0.2, 3))
plot(net, edge.arrow.size=0.2, vertex.size=4, vertex.color=V(net)$color, vertex.label=NA, layout=layout_with_kk)
```
  
  
## 1.5 Node size

We will use the indegree value to assign the node size. First we will calculate indegree using the function `degree()` with the argument `mode = "in"`, and we will also get rid of the loops.  

```{r}
#Calculate the indegree
#mode lets you choose between "in", "out" and "all" for the type of degree
#loops=FALSE removes internal farm shipments
V(net)$indegree <- degree(net,v=V(net),mode=c("in"), loops=FALSE)
#plot the network by adding the vertex.size argument
plot(net, edge.arrow.size=0.2, vertex.size=V(net)$indegree, vertex.color=V(net)$color, vertex.label=NA, layout=layout_with_kk)
```

As you can see, there is one node that receives a lot of movements and you cant really see the edges of the network.  

**Excercise:** since the node size is too large, rescale the values of the node size so you can see the eddges of the network.  

We can also obtain the outdegree using the argument `mode = "out"` side the function `degree()`.  

```{r}
#give weight to node: outdegree#
#same steps with outdegree#
V(net)$outdegree <- degree(net,v=V(net),mode=c("out"), loops=FALSE)
```

**Excercise:** Now change the node size using the outdegree value.  

```{r}
# Now we will calculate the betweeness and use the network to visualize it
# We use the argument directed=T to consider the directionality of the movement
bet <- betweenness(net, v=V(net), directed=TRUE)
# We make sure that we reascale it so we can see the edges
V(net)$size <- scales::rescale(bet, c(5, 15))
plot(net, edge.arrow.size=0.2, vertex.size=V(net)$size, vertex.color=V(net)$color, vertex.label=NA, layout=layout_with_kk)
```

# 2 Using other packages to plot the network

Besides `igraph`, there are other packages to plot the network in R. The package `network()` provides almost the same functions that igraph for plotting, but igraph has more functions for network statistics. To switch between the formats, there is a package calls `intergraph()` which provides functions to going from an igraph object to a network object and the other way.    

## 2.1 network package 
Here we will use the package `network()` to plot a network with the names scaled to the number of neighbors.  

```{r fig.height=10, fig.width=10}
# First we obtain the number of neighbors for each node
V(net)$nb <- neighborhood.size(net)
# then we switch the format to the network package
G <- intergraph::asNetwork(net)
# Now we plot the network
plot(G, # The name of the network
     displaylabels = T, # We want to show labels
     label.cex = rescale(V(net)$nb, to = c(0.001, 1.5)),# We use the number of neighbors to scale the size
     vertex.cex = 0.001, # Reduce the node size so we dont see it
     arrowhead.cex = 0.001, # reduce the arrowhead size
     label.pos = 5, # the position of label (5 = center)
     edge.col = "lightgreen") # Edge color
```

## 2.2 ggraph

Other option for plotting is the package `ggraph` which uses the same syntax than `ggplot`.  

```{r}
net %>%
  as_tbl_graph() %>%
  ggraph(layout = 'kk') +
  geom_edge_link(aes(alpha = pigs.moved)) + # change the transparency based on number of animals moved
  geom_node_point(aes(col = farm_type, size = indegree)) + # size based on indegree
  theme_void()
```

Now we will add the network statistics to the data.frame so we can use them later.  

```{r}
node <- node %>%
  mutate(indegree = V(net)$indegree, outdegree = V(net)$outdegree, nb = V(net)$nb, Bet = bet)
```


Finally we will export the network as a RDS file so we can load it in another session.  

```{r eval = F}
# we can export the network as a RDS object using the package readr
readr::write_rds(net, "Data/Outputs/net.rds")
# We will also export the edges with the distance and coordinates
write.csv(edge, "Data/Outputs/edge.csv", row.names = F)
write.csv(node, "Data/Outputs/node.csv", row.names = F)
```