---
title: "Lab 1"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

In this lab we will start making our first network objects, obtain network statistics and some visualization of the data.  


# Network visualization

## Loading the data and visualizing the network

```{r message = F, warning = F}
# Libraries we will use:
library(dplyr) # for data manipulation
library(ggplot2) # For making figures
library(ggraph) # For ploting the networks
library(igraph) # for network analysis
library(tidygraph) # tidyverse friendly network analysis
# Loading the data from the STNet package
node <- STNet::SwinePrem %>% 
  mutate(id = as.character(id))

edge <- STNet::SwineMov %>% 
  mutate(id_orig = as.character(id_orig), id_dest = as.character(id_dest))
```


The data sets we will be using are:
  
  - `node` which includes the information of the farms 
  - `edge` which includes the information for the movements between the farms.  

We use the function `as_tbl_graph()` to create the network from a data.frame. This function takes as arguments:`x`  which requires the contacts data and assumes that contacts are directed. If we print the results, we will see a object from the class `tbl_graph`, which we will explain:

```{r create the network}
net <- as_tbl_graph(edge)

net
```

The results of printing the object will show us the number of nodes and edges, we can see that our network has 40 nodes and 1611 edges. Then we can also see the attributes for the nodes (which in this case there is only the name) and the edges (which for this example has from, to, date, pigs.moved, type_orig, and type_dest).  
  
We can treat this object like two data frames that are joined by an key or identification variable. Whenever we want to access one of the data frames to modify it, we can use either the function `activate()` or when using the pipes we can use `%N>%` to call the nodes or `%E>%` to call the edges. In the next example we will add the rest of the node information to our `tbl_graph`

```{r}
net <- net %N>% # <- Notice we are including 'N' inside our pipe to specify we want to access the nodes
  left_join(node, by = c('name' = 'id')) # Now we join to the node dataframe to include other variables

# When we print our object again we will see the new variables:
net
```

To visualize the network we can use the function `plot()`.  

```{r}
#plot network#
#for help look look for: ?plot.igraph#
plot(net)
```

This figure looks a bit messy, to make it more informative we can visualize dome of the parameters. We use the arguments `edge.arrow.size` to adjust the size of the arrowhead, `vertex.size` to adjust the node size, and `vertex.label` to remove the names of the nodes.  

```{r}
#make it clearer: adjust size, remove labels#
plot(net, # Our network object
     edge.arrow.size=0.2, # define the arrow size
     vertex.size=4,  # the sie of the node
     vertex.label=NA) # We remove the name of the nodes
```

**Exercise:** Change the size of the nodes and the arrows to make a different graph.  

## Using ggraph

We can use base R to make our figures, but since we will be focusing on ggplot2, we will use the library ggraph which is based in ggplot2.  

### Changing the node color

We can get specific attributes from the `tbl_graph` inside the `aes()` argument, similar to what we would do for ggplot2. In the following plot, we will set the node color to the type of farm.  

```{r}
ggraph(graph = net) + # First we call the ggraph and specify our network
  geom_edge_link() + # Add the edges
  geom_node_point(aes(color = farm_type), size = 3) # add the nodes
```


## Layouts

We can change the position of the nodes using the argument `layout`. If we don't specify any layout, ggraph will automatically set the layout 'stress'. Force directed layouts place the nodes in fixed places based on various criteria. The layouts 'nicely' and 'kk' force the network in clear layouts, based on the position of the nodes in the network in terms of connectivity.  They also attempt to keep the distance between connected nodes fixed.

```{r}
# Plot with layout nicely
ggraph(net, layout = 'nicely') +
  geom_edge_link() + # Add the edges
  geom_node_point(aes(color = farm_type), size = 3) + # add the nodes
  labs(title = 'Layout nicely')

# Plot with layout kk
ggraph(net, layout = 'kk') +
  geom_edge_link() + # Add the edges
  geom_node_point(aes(color = farm_type), size = 3) +
  labs(title = 'Layout KK')
```

You can read more about the different layours available [HERE](https://ggraph.data-imaginist.com/articles/Layouts.html)

Now we will use different layouts to see how the network changes.  

**Excercise:** Try different layouts: 'fr', 'lgl', 'graphopt'.    


## Size of the edge

```{r}
ggraph(net, layout = 'kk') +
  geom_edge_link(aes(width = pigs.moved)) + # Add the edges
  geom_node_point(aes(color = farm_type), size = 3) +
  scale_edge_width(range = c(0.01, 0.9))
```
  
  
## Node size

We will use the indegree value to assign the node size. First we will calculate indegree using the function `centrality_degree()` with the argument `mode = "in"`, and we will also get rid of the loops with the argument `loops = F`.    

```{r}
net <- net %>% 
  mutate(indegree = centrality_degree(mode = 'in', loops = F))

ggraph(net, layout = 'kk') +
  geom_edge_link(aes(width = pigs.moved)) + # Add the edges
  geom_node_point(aes(color = farm_type, size = indegree)) +
  scale_edge_width(range = c(0.01, 0.9))
```

We can use other network centrality measures such as outdegree, closeness, betwenness, among others.

```{r}
net <- net %>% 
  mutate(outdegree  = centrality_degree(mode = 'out', loops = F),
         closeness = centrality_closeness(),
         betweenness = centrality_betweenness(),
         Nbs = neighborhood.size(graph = .)
         )

net
```


**Excercise:** explore plotting the network using the different centrality measures to change its color or size.  


# Using other packages to plot the network

## Interactive visualization using networkD3

The package `networkD3` offer some interesting visualization options. This package uses javascript to create interactive visualizations so we need to start our indexing at 0 and not at 1 for the edges. for this we will substract 1 to the edges indexing

```{r}
# Loading the library
library(networkD3)
# SUbstract 1 from the indexing of the nodes
edge_s <- edge %>%
  count(id_orig, id_dest) %>%
  mutate(id_orig = as.numeric(id_orig) - 1, id_dest = as.numeric(id_dest) - 1)

# Plot a simple network
simpleNetwork(edge_s)
```

We can also add more information to our network such as the name, type of node and size of the edges.  

```{r}
# Plot
forceNetwork(Links = edge_s, Nodes = node,
            Source = "id_orig", Target = "id_dest",
            Value = "n", NodeID = "name",
            Group = "farm_type", opacity = 0.8)
```

Another interesting visualization from this package are the Sankey Networks. THis type of visualizations can be very helpful to se the flow of the movements. The height of the rectangles will represent the indegree, and the shaded line between two nodes represent the number of movements between a pair of nodes

```{r}
# Plot
sankeyNetwork(Links = edge_s, Nodes = node, Source = "id_orig",
             Target = "id_dest", Value = "n", NodeID = "name",
             fontSize = 12, nodeWidth = 30)
```
  
  
# Calculating euclidean distance

We can also use the spatial location of the nodes to calculate the distance between a movement. To represent this distance we use euclidean distance, which is the distance using a straight line between two locations.  
We will use the package `raster` for this part, so make sure its installed.  

```{r warning=F, message=F}
library(raster)
# First we need to include the information of origins and destination
edge <- edge %>%
  left_join(node[c("id", "lat", "long")], by = c("id_orig" = "id")) %>%
  rename(O_Lat = lat, O_Long = long) %>%
  left_join(node[c("id", "lat", "long")], by = c("id_dest" = "id")) %>%
  rename(D_Lat = lat, D_Long = long)
# Now we calculate the distance
edge$dist <- pointDistance(p1 = edge[c("O_Long", "O_Lat")], # Origin coordinates
                           p2 = edge[c("D_Long", "D_Lat")], # Destination coordinates
                           lonlat = T)

# We convert from m to km
summary(edge$dist/1000)
# We can see the quantiles:
quantile(edge$dist/1000, c(0.05,0.95))
# Lets look the distribution of distances
hist(edge$dist/1000, 
     xlim=range(0,500),
     breaks=20,  
     main = "Distribution of euclidean distance (Km)", xlab="distance (km)")

boxplot(edge$dist/1000,
        ylim=c(0,500),
        main = "Boxlot of euclidean distance (Km)", ylab="distance (km)")
```



```{r}
node <- net %N>% 
  data.frame()
```

Now we will add the network statistics to the data.frame and export the objects so we can load it in another session.  

```{r eval = F}
# we can export the network as a RDS object using the package readr
readr::write_rds(net, "Data/Outputs/net.rds")
# We will also export the edges with the distance and coordinates
write.csv(edge, "Data/Outputs/edge.csv", row.names = F)
write.csv(node, "Data/Outputs/node.csv", row.names = F)
```
  

____________________________________________________________________
