---
title: "Lab 2"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

# Community detection

First we will have to load the data. Remember that on the previous lab we exported a dataset at the end, we will continue using this.  

```{r}
# If you are starting a new session, load the files and libraries again 
net <- readRDS("Data/Outputs/net.rds")
# Load the data
node <- read.csv("Data/Outputs/node.csv")
edge <- read.csv("Data/Outputs/edge.csv")
# libraries we will use.
library(igraph)
library(dplyr)
library(sf)
```

Now we will use the walktrap algorithm for community detection in the observed network. For this we use the function `cluster_walktrap()`.   

```{r}
# First we need to simplify the network
c<-simplify(net)
# Then we can use the function
wc <- cluster_walktrap(c)
# To identify the communities we use:
membership(wc)
```
  
We can put this communities on a plot using the base `plot()` function. For this we will need to include in the arguments our simplified network and the communities we detected using the walktrap algorithm.  

```{r}
# We can plot the network and the communities.
plot(wc, c, edge.arrow.size=.2,vertex.size=4,vertex.label=NA)
```


# Spatial representation of the network

Now we will use the network we created and the spatial location of our farms to see the movements on a map.  
We will be using the `sf` package to manipulate the spatial objects, and the `ggplot2` package for visualization.  
In the `STNet` package there is a spatial polygons data, which includes the counties in the state of Iowa.  

```{r}
# Loading the packages
library(sf) # Package for spatial objects
library(ggplot2) # package for plots

# We load the spatial object from the package STNet
iowa <- st_read(system.file("data/Io.shp", package = "STNet"))

# plot map using sf
plot(iowa$geometry)
```

Next we will transform the nodes as a spatial points object, for this we use the function `st_to_sf()` and we need to specify the names of the columns that have the spatial coordinates.    

```{r Transform nodes to spatial}
NodeSp <- node %>% # This is our node data.frame
  st_as_sf(coords = c("long", "lat"), # Variables for the coordinates
           crs = st_crs(iowa)) # This is the CRS we are using
```


## Plotting our map

One of the nice things of ggplot is that we can create a map and store it in an object and later we can keep adding layers to this map. So first we will create a map of the state.    

```{r}
map <- ggplot() +
  geom_sf(data = iowa, # name of the spatial dataset
          color="grey20", # color of the shape border
          fill="white", # fill of the shape
          size=0.4) + # width of the border
  theme_void() # This is a theme form ggplot
```


## Plot the nodes

Once we have the base map of the state, we can add the spatial points data we created previously. We can specify the size of the points using a variable.  

```{r}
map + geom_sf(data = NodeSp, # name of our data
              aes(color = farm_type, # we color the nodes by farm type
                  size = indegree)) +
  ggtitle("Farms and their indegree") # the title of our plot
```

**Excercise:** Make the same plot, but make the size of the nodes relative to outdegree



## Adding the euclidean contacts

In the previous lab we calculated the euclidean distance between each pair of farms involved in a movement. Here we will visualize those movements.  

```{r}
# The function geom_segment adds a straight line between two coordinates:
map +
  geom_segment(data=edge, 
               aes(x=O_Long, y=O_Lat, # this is where the line starts
                   xend=D_Long, yend=D_Lat)) # this is where it ends

# We can add the information of the type of movement to change the color of the line and the number of animals for the transparency
map +
  geom_segment(data=edge, 
               aes(x=O_Long, y=O_Lat,
                   xend=D_Long, yend=D_Lat, 
                   color=type_orig,
                   alpha = pigs.moved))
```


## Putting everything together

Now we will add both the farm locations and the direction of the movements between the farms on a map.  

```{r}
#plot nodes & edges - add both commands geom_segment and geom_point#
map +
  geom_segment(data=edge, 
               aes(x=O_Long, y=O_Lat,
                   xend=D_Long, yend=D_Lat,
                   alpha = pigs.moved), 
               show.legend=F) +
  geom_sf(data = NodeSp, 
          aes(color = farm_type, 
              size = indegree), show.legend = "point")
```


## Subsetting the data

Sometimes we are interested in a particular type of movements. We can subset this using the dplyr functions such as `filter()`. In the next plot we will select only the movements that comes from sow farms.  

```{r}
# plot movements from sow farms only
map +
  edge %>%
  filter(type_orig == "sow farm") %>%
  geom_segment(data = ., 
               aes(x=O_Long, y=O_Lat, 
                   xend=D_Long, yend=D_Lat, 
                   color = type_orig), show.legend = F) +
  geom_sf(data = NodeSp, 
          aes(color = farm_type), size=3, show.legend = "point")
```
  
  We can be even more specific and filter the movements from sow farms that are directed to GDU.  
We will also add at the end the function `ggplotly()` from the package `plotly` to obtain a map were we can zoom and hover over some features to obtain more information.  

  
```{r}
# We store the map of movements between GDU to sow farm
m <- map +
  edge %>%
  filter(type_orig == 'GDU', type_dest == "sow farm") %>%
  geom_segment(data = ., aes(x=O_Long, y=O_Lat,
                             xend=D_Long, yend=D_Lat, 
                             color = type_orig), show.legend = F) +
  geom_sf(data = NodeSp, 
          aes(color = farm_type),
          size=3, show.legend = "point") +
  ggtitle("GDU to Sow farm Movments")

# We use the function from plotly to transform our map into n interactive map.  
library(plotly)

ggplotly(m)
```


# Kernel density map

Like we just saw, visualizing the movements can be challenging, one approach to do this is using a kernel density map. The idea behind this is to extrapolate values in a continuous surface, but here we are just interested in the visualization of the values, not so much in the interpolation of our values.  
We will use the package `KernSmooth` for this, so make sure you ahve it installed.  
First we will define a function to automate the process:  

```{r}
library(KernSmooth)
library(raster)
# we will create a function to create a density raster:
processRaster <- function(x, b, shp, res = c(200, 200)) {
  est <- bkde2D( # we use the function bkde2D to obtain our values
    x, # This will be our dataset
    bandwidth = c(b, b), # The bandwidth we define
    gridsize = res, # the resolution level we want
    range.x = list(extent(shp)[c(1, 2)], extent(shp)[c(3:4)])
  )
  
  # Add the results to a raster
  r <- raster(list(
    x = est$x1,
    y = est$x2,
    z = est$fhat
  )) %>%
    `projection<-`(st_crs(iowa)) %>% # set the CRS
    `extent<-`(extent(iowa)) %>% # set the extent
    crop(., iowa) %>% # crop the raster to the area
    mask(., iowa) # crop the raster to the stat shape
  
  return(r)
}
```

Now lets use our function for the data.  

```{r}
# Obtain the estimated kernel with bandwidth 2km
Erc <- processRaster(x = edge[,c("O_Long", "O_Lat")], # we want the outgoing only
                     b = 2, # we choose a bandwidth of 2
                     shp = iowa) # we set our extent 

# plot the raster and the map
plot(Erc)
plot(iowa$geometry, col=NA, border = "grey80", add = T)
```

We might have used a very large bandwidth in the previous plot, lets try with a smaller size.  

```{r}
# Using a different bandwidth
Erc <- processRaster(x = edge[,c("O_Long", "O_Lat")],
                     b = 0.1,
                     shp = iowa)

# plot the raster and the map
plot(Erc)
plot(iowa$geometry, col=NA, border = "grey80", add = T)
```

**Exercise:** Create a kernel density map for the incoming movements.  

___________________________________________________

# Optional: More on interactive maps. 

If you are interested in more about interactive maps `plotly` also has option for using background maps, but for this you need to get a public [Mapbox access token](https://www.mapbox.com/studio), which is free, but requires registration. Some great resources for more information:  
  
  - https://plotly-r.com/maps.html
  - https://plotly.com/r/maps/

Here I provide an example of the kind of maps you can get using mapbox, but this **This code will not run unless you use your own API Key**
```{r}
# Sys.setenv('MAPBOX_TOKEN' = yourKey)

plot_mapbox() %>%
  add_segments(
    data = group_by(edge, id_orig, id_dest),
    x = ~O_Long, xend = ~D_Long,
    y = ~O_Lat, yend = ~D_Lat,
    alpha = 0.1, size = I(1), hoverinfo = "none"
  )   %>%
  add_markers(data = node,
    x = ~long, y = ~lat, size = ~indegree, text = ~id,
    split = ~farm_type,
    hoverinfo = "text"
  ) %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom = 6,
      center = list(lat = 42, lon = -93)
    ))
```
