---
title: "Lab 3"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

In this section we will use a different dataset which consist on GPS locations of 3 different species.  
We will define our contact network based on proximity between two animals from the GPS location records. 
  
# Exploratory analysis
```{r include=FALSE}
knitr::opts_chunk$set(message = F)
```


```{r message = F}
# libraries we will use
library(dplyr)
library(sf)
library(sp)
library(ggplot2)
# We get the data from the STNet package
GPSc <- STNet::GPSc
```

First we will see how many locations were recorded during the observation period by species:  

```{r}
GPSc %>%
  group_by(species_type) %>%
  summarise(N = n())
```

Now lets create a dataset for the nodes.  

```{r}
Nodes <- GPSc %>%
  distinct(CollarID, species_type)
```

## Temporal visualization

Now lets see how does the GPS records varies during the observation period:  

```{r}
Daily <- GPSc %>%
  count(Date, species_type)

ggplot() +
  geom_line(data = Daily, aes(x = Date, y = n, color = species_type), size = 1) +
  theme_minimal()
```


## Spatial distribution of the records

First we will tronsform our data into a spatial points object.  
```{r}
GPScSP <- GPSc %>%
  st_as_sf(coords = c("X", "Y"), crs = st_crs(4326)) %>%
  st_transform(st_crs(32615))
```

Now we use `geom_sf()` to see the locations on a map.

```{r}
ggplot() +
  geom_sf(data = GPScSP, aes(color = species_type), show.legend = "point") +
  theme_minimal()
```

It looks like the 3 different species have movement ranges where they coincide. To identify where are the species coinciding more often we will use a hexagonal grid.  


# Species range maps

The `STNet` package includes a function called `HexGrid`, which creates a hexagonal grid on a spatial field. Here we will use the extent of our points to create the grid.  

```{r}
# First we create a field that represents the study area:
Border <- as(raster::extent(GPScSP), "SpatialPolygons") %>%
  st_as_sf()
# We define the CRS
st_crs(Border) <- st_crs(GPScSP)
```

```{r}
# Load the STNet library:
library(STNet)
# We use the HexGrid function:
BorderHex <- HexGrid(cellsize = 500, Shp = Border)
# Plot the grid:
plot(BorderHex$geometry)
```

Next we wull use another function from the `STNet` package that counts the number of points per hexagonal cell. The arguments we need are:  
  
  - `Hex`, the hexagonal grid we created.   
  - `Points` The locations that we want to count. for this function is also necesary to have installed the package `sp`, so make sure its installed and loaded.  
  
Now lets plot the movement ranges for the species on the observed period:

```{r}
# For cattle:
GPScSP %>%
  filter(species_type == "cattle") %>%
  HexMap(Hex = BorderHex, .) %>%
  ggplot(., aes(fill = N)) +
  geom_sf() +
  scale_fill_gradient(low="white", high="red") +
  ggtitle("Cattle Range") +
  theme_void()
```

**Exercise:** Create the range maps for the other 2 species (deer y pig) and use different colors.  

```{r eval=FALSE, include=FALSE}
GPScSP %>%
  filter(species_type == "deer") %>%
  HexMap(Hex = BorderHex, .) %>%
  ggplot(., aes(fill = N)) +
  geom_sf() +
  scale_fill_gradient(low="white", high="darkgreen") +
  ggtitle("Deer Range") +
  theme_void()
```

```{r eval=FALSE, include=FALSE}
GPScSP %>%
  filter(species_type == "pig") %>%
  HexMap(Hex = BorderHex, .) %>%
  ggplot(., aes(fill = N)) +
  geom_sf() +
  scale_fill_gradient(low="white", high="darkblue") +
  ggtitle("Pig Range") +
  theme_minimal()
```
  
  

# Defining the network:

## Defining the contacts:

Now we will define our network. We are interested in all the contacts that happened between animals at a distance of < 1 m.  

This function might take some time, if you want to skip this part, the data is available inside the `STNet` package.  

```{r eval = F}
# Lets change the names to test the function:
colnames(GPSc)[c(1, 4, 6, 7)] <- c("id", "date", "Longitude", "Latitude")
# Run the function
Edges <- CreateNetwork(DF = GPSc[,], # This is our data
                       DTh = 1, # TTthe distance threshold
                       DateTime = "date", # Name of the variable that indicates the date and time
                       ID = "id", # name of variable for id
                       coords = c("Longitude", "Latitude") # Name of variable for coordinated
                       ) # 14,869
```

To load the data directly from the `STNet` package we can use the function:  

```{r}
Edges <- STNet::.
```
This dataset includes:  
  
  - *Distance*: distance between the recorded contact.  
  - *DateTime*: The time when the contact happened.  
  - *Var1* and *Var2*: nodes involved in the contact. Remember that since these contacts are registerd from the location of the animals, we don't know who started the contact, so it is not directed.  
  - *X.x*, *Y.x*, *X.y*, *Y.y*: the spatial location of the nodes at that time.  
  
### Contacts between species:

We can also detect areas where the contacts between species are happening.  
For this we will create a subset of the data that includes only contacts between different species.  

```{r}
# Add variable for species
Edges <- Edges %>%
  left_join(Nodes, by = c("Var1" = "CollarID")) %>% #Var1 one of the nodes
  rename(Sp1 = species_type) %>%
  left_join(Nodes, by = c("Var2" = "CollarID")) %>% # Var 2 is the other node
  rename(Sp2 = species_type)

InterSp <- Edges %>% 
  filter(Sp1 != Sp2) %>% # select contacts where the species are different
  data.frame()
```


```{r}
InterSp_sp <- InterSp %>%
  dplyr::select(DateTime, Var1, Var2, Sp1, Sp2, X.x, Y.x) %>%
  st_as_sf(coords = c("X.x", "Y.x"), crs = st_crs("+init=EPSG:4326")) %>%
  st_transform(st_crs(32615))
  
IspH <- HexMap(BorderHex, Points = InterSp_sp)

ggplot(IspH, aes(fill = N)) +
  geom_sf() +
  scale_fill_gradient(low="white", high="gold3") +
  ggtitle("Contacts between species") +
  theme_void()
```


## Define the nodes:

We previously defined the nodes, we can see that there are 36 animals in our network.

```{r}
Nodes %>%
  count(species_type)
```

  
  
## Creating the network:

```{r}
# Libraries we will use:
library(igraph)
library(scales)
# Create the network:
G1 <- graph_from_data_frame(Edges[c("Var1", "Var2")], vertices = Nodes, directed = F)
# Create a variable to count the number of contact:
E(G1)$w <- 1
# Simplifying the network:
G1s <- simplify(G1, # Name of the network
                remove.multiple = T, # Remove multiple contacts
                edge.attr.comb = list(w = "sum")) # Function to do with the multiple contacts

# calculate the degree (since its not directed we can not define indegree or outdegree)
V(G1)$degree <- degree(G1, mode = "all")

plot(G1s, 
     edge.color = alpha("black", rescale(E(G1s)$w, c(0.01, 1))), # use number of contacts for transparency
     layout = layout_with_kk, 
     vertex.label = NA, 
     vertex.size = rescale(V(G1)$degree, c(1, 15)))
```


## Networks by species: 
  
To create a subset of our network we can use the function `induced_graph`, where we need to specify the network that we will extract from and the ids of the nodes, for this we use the function `which()` and specify that we watn the deer only.  

```{r}
VIDs <- which(V(G1)$species_type == "deer")
DeerG <- induced_subgraph(graph = G1s, vids = VIDs)


plot(DeerG, 
     edge.color = alpha("black", rescale(E(DeerG)$w, c(0.01, 1))),
     layout = layout_with_kk, 
     vertex.label = NA, 
     vertex.size = rescale(V(G1)$degree[VIDs], c(1, 15)), 
     vertex.color = "darkgreen")
```

**Exercise:** create the network for the other 2 species

___________________________________________________
