---
title: "Lab 4"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r echo=F}
knitr::opts_chunk$set(message=F, warning =F)
```



These are the libraries that we will use in this lab:


```{r }
# install.packages("igraph")
# install.packages("maps")
# install.packages("raster")
# install.packages("rgdal")
# install.packages("ggplot2")
# devtools::install_github("GIScience/openrouteservice-r")
# install.packages("sp")
# install.packages("maptools")
# install.packages("spacetime")
# install.packages("plotKML")
# install.packages("scatterplot3d")
```

Let's start by importing the data for edges, node attributes, and the shapefile of the area of study.

```{r  echo=FALSE}
colpal <- RColorBrewer::brewer.pal(4, "Dark2")
library(dplyr)
library(sf)
# Load data of movements and nodes 
net <- STNet::SwineMov
node_attrib <- STNet::SwinePrem
# Format id as character to use it later
node_attrib$idch <- as.character(node_attrib$id)
# Load the spatial object
IO <- st_read(system.file("data/Io.shp", package = "STNet"), quiet = T)
```

We will now use the `graph.data.frame` function to create a directed network.

```{r}
library(igraph)
g1 <- graph.data.frame(net, directed = T)
```

As we have seen earlier, we can add network attributes for both vertices `V ()` or edges `E ()`. In the next section we will add some attributes related with the location of farms and farm type.

```{r}
# We match the identifiers in the network with the data table 
poss<-match(V(g1)$name,node_attrib$idch)
# Add the attributes to the network
V(g1)$lat <- node_attrib$lat[poss]
V(g1)$long <- node_attrib$long[poss]
V(g1)$type <- as.character(node_attrib$farm_type[poss])
V(g1)$fullname <- as.character(node_attrib$name[poss])
# check the correct assignation of the attributes in the object

summary(g1)
```

# Introduction to dynamic networks

The `networkDynamic` package has functions that allow to analyze and visualize a network and incorporate temporality to it.
To incorporate a temporal window, we need to specify the start time (onset) and end time (termini). For the purposes of this example we will assume that the contact time is one week.

```{r echo=FALSE}
# Load libraries
## For dynamic networks
library(networkDynamic)
# To visualize and get network statistics
library(ndtv)
library(tsna)
# Format the variable as date:
net$date <- as.Date(net$date, format = '%m/%d/%y')
# Create a new variable for year from the date
net$year<-as.numeric(strftime(net$date,format = "%Y"))
# Create a new variable to show the week of the year
net$w<-as.numeric(strftime(net$date,format = "%V"))
```

In this analysis we will use a one week resolution. Since we have no information on the duration of the contact, we will assume that the duration is one week. As we have multiple years, we need to number the weeks from the first week in which we have data to the last week with data. For this, we will use an offset that will add 53 weeks or 106 depending on the year.

```{r}
# Create the onset starting from 2015, 
# Use ifelse inside another ifelse to determinate the offset
net$w_onset<-as.numeric(ifelse(net$year==2015 #First condition
                               ,net$w # If the condition is true, only the number of week, if not, we use other ifelse 
                               ,ifelse(net$year==2016 # Second condition
                                       ,net$w+53 # If the condition is true, an offset of 53 weeks is added, 106 weeks otherwise
                                       ,net$w+106)))

# Now we state that the duration of the contact is one week 
net$w_termini<-net$w_onset+1
```

Now we will create a dynamic network with this information.
The `networkDynamic` function is able to reproduce the data in different formats. Here we will use a `data.frame` format. The function will read the data and will expect a list of variables following this order: `head`, `tail`, `onset` and `termini`, where:
`head` = Origins.
`tail` = Destination.
`onset` = Start of contact.
`termini` = End of contact.

```{r}
n1 <- networkDynamic(edge.spells = net[c("w_onset", "w_termini", "id_dest", "id_orig")])
```

Now we can add attributes to the data. These attributes can be static or change during the period. 
**Note:**  the functions `set.edge.attribute()` and `set.vertex.attribute()` from the package network can be masked by other functions with the same name that are present in other packages, such as the `igraph` package. If we have already loaded these other packages with functions in conflict, we will need to specify which package contains the function that we want to use to prevent wrong calls (i.e. by typing `network::set.edge.attribute()`).

```{r}
set.edge.attribute(n1, #Name of the network
                   "pigs.moved",  # Name of the attribute
                   net$pigs.moved) # Values
```

We can then just add the unchanging vertex attribute data

```{r}
set.vertex.attribute(n1,"name", as.character(node_attrib$name))
set.vertex.attribute(n1,"lat",node_attrib$lat)
set.vertex.attribute(n1,"long",node_attrib$long)
set.vertex.attribute(n1,"type",as.character(node_attrib$farm_type))
```

Once we have the node attributes, we can print them by calling the `networkDynamic` object that we have just created 

```{r}
n1
```

This object summarizes the network including the period and number of vertices and edges

```{r}
list.edge.attributes(n1)
```

```{r}
list.vertex.attributes(n1)
```

If we want to see the values of these attributes, we can use  `get.vertex.attribute()`:

```{r}
get.vertex.attribute(n1,"vertex.names")
```

```{r}
get.vertex.attribute(n1,"type")
```

We can also obtain more information of the network, such as the activity by using `get.vertex.activity()` or `get.edge.activity()`

```{r}
summary(get.vertex.activity(n1, as.spellList=TRUE))
```

```{r}
summary(get.edge.activity(n1, as.spellList=TRUE))
```

# Temporal statistics

We can evaluate several measures at the graph level, vertex level or traces of the period of study of our network using the function `tSnaStats()`.

This function uses arguments to include the `networkDynamic` object that we have created and to specify the function with the desired statistic that we want to apply (through the argument "snafun="). For example, if we want to calculate the density of the network in different periods, we will do this:

```{r}
tSnaStats(n1, snafun = "gden") %>% # "gden" to show graph density
  plot(col = colpal[2], lwd = 2, main = "Network density during the study period")
```

In the previous plot we can see that the activity of movements is bigger at the beginning of the time series and slowly decreases  over time until reaching approximately zero.

The function `tSnaStats()` also allows us to use arguments to define specific periods of time from which we want to calculate the statistics

```{r}
tSnaStats(n1, snafun = "grecip", start = 30, end = 60) %>% 
  plot(main = "Network Reciprocity during the interval 30-60", col = colpal[3], lwd = 2)
```

We can also see node-level statistics (e.g. betweenness) using the next function.

```{r}
apply(tSnaStats(n1, snafun = "betweenness"), 1, mean) %>% plot(type = "l", lwd = 2, col = colpal[4])
```

# Plotting the network

We can use the function `plot()` to see the network object, but we can only see the static network because we are not defining any time interval

```{r}
plot(n1,mode="kamadakawai") 
```

If we want to see a specific period of time, we need to define it. For this, we can use the function `network.extract()` to extract a specific period in the network and then use `plot()` to show this extract. 

```{r}
# Plot an interval of time
n1 %>% 
  network.extract(onset = 1, terminus = 60) %>%
  plot()
```

```{r}
# Plot a specific timepoint and remove isolates
n1 %>% 
  network.extract(at = 35) %>%
  plot(displayisolates = F)
```

Now we are going to create a plot with multiple periods of time

```{r}
# First we create a function to plot the network extract we want:
plot.extract <- function(n, onset, terminus){
  n %>% # The input network
    network.extract(onset = onset, terminus = terminus) %>% #Extract the time interval
    plot(vertex.cex = 1.5, main = paste0("t", onset, "-", "t", terminus)) #Plot the network
}

# Then we define out time intervals
O <- seq(from = 25, to = 95, by = 10)
# We set the layout we want for the plot:
par(mfrow = c(2,4), mar = c(0,0,1,0))
# We use a foor loop to plot the network at the defined time intervals.
for(i in O){
  plot.extract(n1, onset = i, terminus = i + 10)
}
```

We can use different layouts to plot the extracts of the network.
Now, we are going to plot the nodes in their spatial location. For this, we will have to make a little modification of our previous function so we can add the coordinates.

```{r}
# We set fixed coordinates from the longitude and latitude
a <- as.matrix(node_attrib[, c("long", "lat")])


# Modify the function:
plot.extract <- function(n, onset, terminus, Bg){
  plot(Bg, col = "grey80", main = paste0("t", onset, "-", "t", terminus))
  n %>% # The input network
    network.extract(onset = onset, terminus = terminus) %>% #Extract the time interval
    plot(vertex.cex = 1, coord=as.matrix(a), new = F) #Plot the network
  
  # plot(network.extract(n1,onset=onset,terminus=terminus),vertex.cex = 1, coord=as.matrix(a), main = paste0("t", onset, "-", "t", terminus))
}


# We set the layout we want for the plot:
par(mfrow = c(2,4), mar = c(0,0,1,0))
# We use a foor loop to plot the network at the defined time intervals.
for(i in O){
  plot.extract(n1, onset = i, terminus = i + 10, Bg = IO$geometry)
}
```

# Temporally reachable sets

The reachability of a farm help us to identify important nodes in the network that may have a key role in the disease spread. Forward reachability ("fwd") refers to all the nodes than can be reached from a given node $v_i$ and backward reachability ("bkwd") refers to those nodes that reach a given node $v_i$.
We can visualize the distribution of the forward and backward reachability using boxplots

```{r}
fwd <- tReach(n1, "fwd")
bwd <- tReach(n1, "bkwd")
par(mar=c(2,4,2,2))
boxplot(fwd, bwd, col = "lightgreen", horizontal = T, yaxt = 'n', main = "Reachability of farms")
axis(2, at = c(1,2), labels = c("Fwd", "Bwd"), las=2, lty = 0)
```

We can also use a dot plot to explore whether any interesting relationship between both types of reachability exists.

```{r}
plot(fwd, bwd, xlab = "Fwd", ylab = "Bwd", pch = 16, col = rgb(0, 155, 50, max=255, alpha=100))
```

Now, let's have a look at the farms with the highest forward reachability:

```{r}
data.frame(id = 1:length(fwd), fwd, bwd) %>% 
  arrange(desc(fwd)) %>% 
  head(10)
```

It looks like node ("id") 7 presents the highest number of reachable farms. 

We can visualize the reachable st using the functions `tpath` and `transmissionTimeline`. Both of them provide the node id and the time when the contact happened. The first function allow us to visualize the nodes spatially and the second one shows them in a nice timeline. Let's start with the first one.

```{r}
P17 <- tPath(n1, 17, direction = "fwd")
plot(P17)
```

We can also visualize the paths in the full network.

```{r}
plotPaths(n1, P17, vertex.col = colpal[1], label.cex = 0.5, vertex.cex = 2)
```

A better way to visualize the paths trought the time period is using the function `transmissionTimeline()`. In this visualization, the x axis represents the time when the event happened and the y axis represents the generation or steps that happened to reach that node

```{r}
transmissionTimeline(P17,jitter=T,
          main='Earliest forward path from vertex 137')
```

We can also see the backward reachability, which represents the set of nodes that reach a specific node over a time period.

```{r}
P17 <- tPath(n1, 17, direction = "bkwd", type = "latest.depart")
plot(P17, edge.col = colpal[3])
```

Now we will create a function to show the paths in a spatial and non-spatial plot.

```{r}
MaPaths <- function(Net, v, direction, Pts, coords, BG){
  par(mfrow = c(1,2), oma = c(0, 0, 2, 0), mar = c(1, 1, 1, 1))
  P1 <- tPath(Net, v = v, direction = direction)
  plot(BG, col = "grey80", axes = F)
  points(Pts[, c("long", "lat")], pch = 16, cex = 0.5, col = rgb(0, 150, 100, max = 255, alpha = 50))
  plot(P1, coord = coords, displaylabels = F, edge.label.col = rgb(0,0,0,0), edge.lwd = 2, vertex.col = "white", edge.col = rgb(0, 200, 50, max=255, alpha=166), new = F)
  
  plot(P1, displaylabels = F,edge.lwd = 2, vertex.col = "white", edge.col = rgb(255, 0, 50, max=255, alpha=166))
  mtext(paste0("Reachable set of farm", v), outer = TRUE, cex = 1.5)
}

MaPaths(Net = n1, v = 17, direction = "fwd", Pts = node_attrib, coords = a, BG = IO$geometry)
```

Another way to visualize the network activity is to show the time points when each of the nodes was active. In the following plot, the upper part represents the presence of the node in the network and the lower part represents the activity of the node in the network.

```{r}
timeline(n1)
```

# Animations

We can also create animations to show the evolution of the contact patterns over time using the function `render.d3movie()`. First we will use the function `compute.animation()` to set the layout and time frame for the animation.

```{r message=FALSE, warning=FALSE,eval=F}
render.d3movie(n1,
               displaylabels=F, 
               launchBrowser = T, 
               vertex.cex = scales::rescale(degree(n1), to = c(1, 5)), filename = "Animation01.html") 
```

If we would like to take specific frames for the animation, we can show them stacked in a plot using the function `timePrism()`.
This visualization method works better for small networks.

```{r message=FALSE}
# Define the intervals
compute.animation(n1, animation.mode = "kamadakawai",
                  slice.par=list(start=30, end=100, interval=1, 
                         aggregate.dur=1, rule='any'))

timePrism(n1, at = c(30, 40, 50, 60), planes = T)
```

________________________________
